Certified Symbolic Finite Transducers: Formalization and Applications to String Analysis

### 2025-09-13 Paper/abstract.tex
abstract and intro

\begin{abstract}
    Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by
    enabling the transformation of input strings into output strings. In many
    practical applications --- including program analysis, string constraint
    solving, and analysis of security-critical sanitizers --- Symbolic FTs 
    (SFTs) and Symbolic FAs (SFAs) are used instead of the explicitly
    represented models. To circumvent to the notorious state-space explosion 
    problem caused by an extremely large alphabet size (e.g. UTF-16), SFTs
    and SFAs allow the representation of the alphabet as an effective boolean
    algebra including finite unions of intervals, as well as 
    SMT-Algebras.
    %can be instantiated with various Boolean algebras, 
    The
    security-critical nature of many of these applications demands trustworthy
    implementations of such systems. To this end, we present the first 
    formalization of SFTs and their most important algorithms in
    Isabelle/HOL. 
    %This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
%
    To evaluate the effectiveness of our formalization, we apply the formalized 
    SFTs to two applications: (1) sanitizers for web applications used for 
    preventing XSS attacks, and (2) string solving, which increasingly
    employs intricate string replacement operations.
    %a string solver that models replacement operations with various semantics, including left-most and replace-all matching. 
    Our experimental results demonstrate that our methods are competitive with
    the existing unverified implementations.

>>>>

\begin{abstract}
    Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by
    enabling the transformation of input strings into output strings. In many
    practical applications --- including program analysis, string constraint
    solving, and analysis of security-critical sanitizers --- Symbolic FTs 
    (SFTs) and Symbolic FAs (SFAs) are used instead of the explicitly
    represented models. To circumvent to the notorious state-space explosion 
    problem caused by an extremely large alphabet size (e.g. Unicode), SFTs
    and SFAs allow the representation of the alphabet as an effective boolean
    algebra including finite unions of intervals, as well as 
    SMT-Algebras.
    %can be instantiated with various Boolean algebras, 
    The
    security-critical nature of many of these applications demands trustworthy
    implementations of such systems. To this end, we present the first 
    formalization of SFTs and their most important algorithms in
    Isabelle/HOL. 
    %This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
%
    To evaluate the effectiveness of our formalization, we apply the formalized 
    SFTs to two applications: (1) sanitizers for web applications used for 
    preventing XSS attacks, and (2) string solving, which increasingly
    employs intricate string replacement operations.
    %a string solver that models replacement operations with various semantics, including left-most and replace-all matching. 
    Our experimental results demonstrate that our methods are competitive with
    the existing unverified implementations.
